---
title: "Some statistics"
author: "Rosie"
date: 2023-10-24
categories: ["R"]
tags: ["R Markdown", "plot", "regression"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, echo=FALSE)
```

```{r, message=FALSE}

library(knitr)
library(flextable)
library(tidyverse)
library(blogdown)

serve_site()
```

# Unsustainable Cost

As a rule of thumb of affordability, no more than 30% of one's gross annual income should go towards housing related expenses. In 2021, a staggering 45% of Toronto renter households spent more than 30% of their income on housing. (See Table \@ref(tab:T1).)

```{r T1, results='asis'}

# Table \@ref(tab:T1)

dat <- read.csv("~/housing data/housingcost_over30percent.csv")
names(dat) <- c("location", "total", "owner", "renter")
#head(dat)

# knitr::kable(dat, 
#       col.names = c("Location", "Total households", "Owner households", "Renter households"), 
#       caption = "Percent of households spending 30% or more of income on shelter cost in primary downtowns by tenure, 2021 (Source: Statistics Canada)") 
  

flextable(dat) %>% 
  add_header_row(values = c("Location", "% of households"),
                  colwidths = c(1, 3)) %>%
  set_header_labels(location = "Location", 
                    total = "Total households", 
                    owner = "Owner households", 
                    renter = "Renter households") %>%
  merge_v(part = "header") %>%
  width(width = c(1.5, 1.6, 1.6, 1.6)) %>%
  set_caption(caption = "Percent of households spending 30% or more of income on shelter cost in primary downtowns by tenure, 2021") %>%
  add_footer_lines("Source: Statistics Canada")

```

The debt to income ratio is an aggregate index which compares the total debt to annual income. Canada's 2023 Q2 ratio is 180.54%. It means for every dollar of disposable income, the average household owes about $1.85 in credit market debt, which includes consumer credit and mortgage and non-mortgage loans. It is not the most intuitive concept but it is useful when looking at historical trends. To nobody's surprise, household debt has been increasing since 1990. Interestingly, the pace of increase was not dramatic from 2010, compared to the period of 2000 to 2010. Source: Household sector credit market summary table, seasonally adjusted estimates.
<https://www150.statcan.gc.ca/t1/tbl1/en/cv.action?pid=3810023801>

```{r}

dat <- read.csv("~/housing data/debt_to_income_ratio.csv", header=FALSE)
head(dat)

dat <- as.data.frame(t(dat))
head(dat)

dat <- dat[-1, ]
rownames(dat) <- NULL
head(dat)

names(dat) <- c("year_qt", "ratio")
head(dat)

dat <- dat %>% 
  separate(year_qt, c("quarter", "year")) %>%
  mutate(year = as.numeric(year), ratio = as.numeric(ratio))

dat_q2 <- dat %>% filter(quarter=="Q2")

ggplot(data=dat_q2, aes(x=year, y=ratio, group=1)) +
  geom_line()+
  geom_point() + 
  scale_x_continuous(breaks = c(seq(1990, 2022, by = 2), 2023)) + 
  scale_y_continuous(breaks = seq(80, 190, by = 10)) +
  labs(x = "Year", y = "Debt-to-income-ratio", caption = "Source: Statistics Canada") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=0.5))
```


